% ************************** Thesis Acknowledgements **************************

\begin{acknowledgements}  

I am grateful to my supervisors Nauzika Kovács and Miroslav Sýkora for their guidance and continuous support during the completion of this work. I owe my gratitude to László Dunai for allowing me to pursue the whim of my interest over the course of my studies. This work would have never been completed without his support and his trust in me. I am indebted to László Gergely Vigh for the valuable discussions and for jointly completed research. I highly appreciate his personal and professional help, which he is always eager to provide even amid his numerous activities.

I am thankful to all members of the Department of Structural Engineering for the friendly and supportive environment, for sharing their knowledge, and for nurturing my engineering insight. Special thanks go to Kitti Gidófalvy, Piroska Mikulásné Fegyó, and Mansour Kachichian for the personal conversations.

I am grateful to the members of Klokner Institute, particularly to those of the Department of Structural Reliability, where I spent a wonderful 10-month research stay. I am especially grateful to Miroslav Sýkora for making my stay in Prague an unforgettable experience, for the invigorating professional conversations, and for his friendship. 

Special thanks to commander-in-chief Mogyi for her professional help and for her companionship in our journey {\tiny \PHcat}.

I am obliged to the countless people of past and present who shaped my world view and who's work and ideas are indispensable and inseparable part of my thinking. I wish I were able to add a grain of sand to the mountain of their amassed knowledge. 

This  work  was partly supported by the Ph.D. Scholarship of the Hungarian State and by the International  Visegrad  Fund  Intra-Visegrad Scholarship  (contract  no.  51401089), both are highly appreciated.

The numerical analyses are mainly completed using Matlab \citep{Matlab2015a}, FERUM \citep{Kiureghian2006}, and R \citep{R2015}. The work and commitment of the developers of these applications are highly appreciated. In the spirit of reproducible and open research all of the related scripts, processed data, and results are available from the author and partially from the repositories of the following GitHub account:\\
\url{https://github.com/rozsasarpi}.

\end{acknowledgements}
